---
title: "p8105_hw1_rkk2139"
author: "Riya Kalra"
output: github_document
---

<font size="5"><b>Problem 1</b></font>

```{r setup, include = FALSE}
library(tidyverse)
library(dplyr)
```
#Load dataset
```{r}
data("penguins", package = "palmerpenguins")
```
The **penguins** dataset contains various variables for different penguin species in the Palmer Archipelago. It includes `r nrow(penguins)` rows and `r ncol(penguins)` columns, and important variables include:

- `species`: species of penguins (Adelie, Chinstrap, Gentoo),
- `island`: the island where the penguins were observed,
- `bill_length_mm`: the length of the penguin's bill in millimeters,
- `bill_depth_mm`: the depth of the penguin's bill in millimeters,
- `flipper_length_mm`: the length of the penguin's flipper in millimeters, and
- `body_mass_g`: the body mass of the penguin in grams.

The mean flipper length across all penguins is `r round(mean(penguins$flipper_length_mm, na.rm = TRUE), 2)` mm.

```{r, warning=FALSE}
# Scatterplot of flipper_length_mm vs bill_length_mm, colored by species
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) +
  geom_point() +
  labs(title = "Flipper Length vs Bill Length by Species",
       x = "Bill Length (mm)",
       y = "Flipper Length (mm)") +
  theme_minimal()
```
  

```{r, warning=FALSE}
# Export the scatterplot to the project directory
ggsave("penguins_flipper_vs_bill.png", width = 6, height = 4)
```

<font size="5"><b>Problem 2</b></font>

```{r}
# Create a data frame
coercion_df <- data.frame(
  numeric_sample = rnorm(10),  # Random sample from a standard Normal distribution
  logical_vector = rnorm(10) > 0,  # Logical vector indicating whether elements are greater than 0
  char_vector = sample(letters, 10),  # Random character vector of length 10
  factor_vector = factor(sample(c("Level1", "Level2", "Level3"), 10, replace = TRUE))  # Factor vector with 3 levels
)

print(coercion_df)

# Try to take the mean of each variable
mean_numeric <- mean(pull(coercion_df, numeric_sample))  # Works, because it's numeric
mean_logical <- mean(pull(coercion_df, logical_vector))  # Works, logical values are treated as 0 (FALSE) or 1 (TRUE)
mean_char <- mean(pull(coercion_df, char_vector))  # Doesn't work, you can't take the mean of characters
mean_factor <- mean(pull(coercion_df, factor_vector))  # Doesn't work, factors are not numeric

mean_numeric  # Should work
mean_logical  # Should work (will be a proportion of TRUEs)
mean_char     # Will throw an error
mean_factor   # Will throw an error

# Convert variables to numeric types
numeric_logical <- as.numeric(coercion_df$logical_vector)  # Converts TRUE to 1, FALSE to 0
numeric_char <- as.numeric(coercion_df$char_vector)  # Will result in an error, cannot convert characters to numeric
numeric_factor <- as.numeric(coercion_df$factor_vector)  # Converts factors to their underlying integer codes

print(numeric_logical)
print(numeric_factor)

# numeric_char will give an error
```


