---
title: "p8105_hw1_rkk2139"
author: "Riya Kalra"
output: github_document
---

<font size="5"><b>Problem 1</b></font>

```{r setup, include = FALSE}
library(tidyverse)
```
#Load dataset
```{r}
data("penguins", package = "palmerpenguins")
```
The **penguins** dataset contains various variables for different penguin species in the Palmer Archipelago. It includes `r nrow(penguins)` rows and `r ncol(penguins)` columns, and important variables include:

- `species`: species of penguins (Adelie, Chinstrap, Gentoo),
- `island`: the island where the penguins were observed,
- `bill_length_mm`: the length of the penguin's bill in millimeters,
- `bill_depth_mm`: the depth of the penguin's bill in millimeters,
- `flipper_length_mm`: the length of the penguin's flipper in millimeters, and
- `body_mass_g`: the body mass of the penguin in grams.

The mean flipper length across all penguins is `r round(mean(penguins$flipper_length_mm, na.rm = TRUE), 2)` mm.

```{r}
# Scatterplot of flipper_length_mm vs bill_length_mm, colored by species
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) +
  geom_point() +
  labs(title = "Flipper Length vs Bill Length by Species",
       x = "Bill Length (mm)",
       y = "Flipper Length (mm)") +
  theme_minimal()
```
  

```{r}
# Export the scatterplot to the project directory
ggsave("penguins_flipper_vs_bill.png", width = 6, height = 4)
```

